Parsing Tweets into Universal Dependencies

### 2017-12-13 paper.tex
Noah

\section{Introduction}
NLP for social media messages is challenging, both because domain
adaptation is required, but also because creating annotated datasets
(e.g., treebanks)
for training and evaluation is hard. 
Pioneering work by \citet{AAAIW113912} 
annotated 7,630 tokens' worth of tweets according to the
phrase-structure conventions of the Penn Treebank
\citep[PTB;][]{Marcus93buildinga}, enabling conversion to Stanford dependencies.
\citet{kong-EtAl:2014:EMNLP2014} further studied the challenges in
annotating tweets and presented a tweet treebank ({\sc Tweebank}),
consisting of 12,149 tokens and largely following conventions
suggested by \citet{schneider-EtAl:2013:LAW7-ID}, fairly close to 
\citet{Yamada03statisticaldependency} dependencies (without labels). 
Both annotation efforts were highly influenced by the PTB, whose guidelines
have good grammatical coverage on newswire. However, when it comes
to informal, unedited, user-generated text, the guidelines may leave
many annotation decisions unspecified.
%Motivation notes:  explain why UD is attractive (and more attractive
%than what previous work used, which was based on YM).  one reason is
%that it might allow sharing of texts annotated in different genres (or
%even languages, cite Waleed).  but universal annotation schemes are
%challenging!  this paper contends with the challenges of
%applying UD to social media texts.

@@

Based on these annotations, we nonetheless designed a pipeline to parse 
raw tweets into universal dependencies. Our pipeline includes: a
bidirectional LSTM (bi-LSTM) tokenizer, a word cluster-enhanced POS
tagger \citep[following][]{owoputi-EtAl:2013:NAACL-HLT}, and a stack LSTM parser
with character-based word representations
\cite{ballesteros-dyer-smith:2015:EMNLP}, which we refer to as our
``baseline'' parser.
To overcome the noise in our annotated data and achieve better performance
without sacrificing computational efficiency, we 
distill a 20-parser ensemble into a single greedy  parser 
\cite{DBLP:journals/corr/HintonVD15}.
We show further  that learning directly from the exploration of the ensemble parser
is more beneficial than learning from the gold standard ``oracle''
transition sequence. Experimental results show that an improvement of more
than 2.6 points in LAS over the
baseline parser can be achieved with our distillation method and it outperforms
other state-of-the-art parsers in both accuracy and speed.

The contributions of this paper include:
\begin{itemize}
\item We study the challenges of annotating tweets in UD (\S\ref{sec:anno})
and created a new tweet treebank ({\sc Tweebank v2}), which includes 
tokenization, part-of-speech tagging, and labeled universal dependencies annotation.
We also characterize the difficulties of creating such annotation.

@@

We first review {\sc Tweebank v1} of \citet{kong-EtAl:2014:EMNLP2014},
the largest Twitter dependency annotation effort to-date
(\S\ref{sec:tweebank}).
Then we introduce the differences in our tokenization
(\S\ref{sec:tok-anno}) and part-of-speech (\S\ref{sec:pos-anno}) (re)annotation with \citet{ICWSM101540} and 
\citet{gimpel-EtAl:2011:ACL-HLT2011}, respectively, on which {\sc Tweebank v1} was built. 
We describe our effort of adapting the
UD conventions to cover tweet-specific constructions (\S\ref{sec:ud-tweet}). 
Finally, we present our process of creating a new tweet treebank, {\sc
  Tweebank v2}, and characterize
the difficulties in reaching consistent annotations (\S\ref{sec:anno-process}).

@@

In the different direction of splitting tokens, UD guidelines also suggest to merge
{\it multi-token words} (e.g. {\it 20 000}) into one single token in some special
cases. We witnessed a small number of tweets that contain multi-token words
(e.g. {\it Y O}, and {\it R E T W E E T}) but didn't combine them for simplification.
Such tokens only account for 0.067\% and we use the UD
  {\it goeswith} relation to resolve these cases in the dependency annotations.

@@

Another important difference follows from the difference in
tokenization.  As discussed in \S\ref{sec:tok-anno}, UD calls for more
aggressive tokenization than that of \citet{ICWSM101540} which opted
out of splitting contractions and possessives. As a consequence of adopting \citet{ICWSM101540}'s tokenization, 
Gimpel et al. introduced new parts of speech for these cases
instead.\footnote{These tags only account for 2.7\% of tokens,
  leading to concerns about data sparseness in tagging and all
  downstream analyses.}  For us, these tokens must be split, but
universal parts of speech can be applied.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth,trim={0 0 11cm 9cm},clip]{graphics/bilstm_tokenizer}
	\caption{The bi-LSTM tokenizer that segments `{\it its gonna be}' into `{\it it s gon na be}'.}\label{fig:tok-model}
\end{figure}
We introduce a new
character-level bidirectional LSTM (bi-LSTM) sequence-labeling model
\cite{DBLP:journals/corr/HuangXY15,ma-hovy:2016:P16-1}
for tokenization.
Our model takes the raw sentence and tags each character in this 
sentence as whether it is the beginning of a word (1 as the beginning and 0 otherwise).
Figure \ref{fig:tok-model} shows our tokenization model.
Space is treated as an input but consistently assigned a special tag \$.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & F1 \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UD pipe v1.2 & 94.7 \\
		\hdashline
		UD pipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}

@@

Part-of-speech tagging for tweets has been extensively studied \cite{ritter-EtAl:2011:EMNLP,gimpel-EtAl:2011:ACL-HLT2011,derczynski-EtAl:2013:RANLP-2013,owoputi-EtAl:2013:NAACL-HLT,gui-EtAl:2017:EMNLP20172}. We therefore opt for the existing POS taggers for tweets instead of developing our own's.
On the annotation scheme designed in \S\ref{sec:pos-anno}, based on UD and adapted for
Twitter, we compared several existing systems which include the Stanford CoreNLP tagger, 
\citet{owoputi-EtAl:2013:NAACL-HLT}'s word cluster enhanced tagger (both in greedy and CRF), and 
also \citet{ma-hovy:2016:P16-1}'s neural network POS tagger 
which has the state-of-the-art performance  on Penn Treebank data.
 %and also two bi-LSTM systems similar
%to \citet{DBLP:journals/corr/HuangXY15} and \citet{lample-EtAl:2016:N16-1}, one using
%word vectors and the other using word and character vectors (see Figure \ref{fig:pos-model}).
All systems were re-trained on the combination of the UD\_English and 
\textsc{Tweebank v2} training sets. We use Twitter-specific glove embeddings released by
\citet{pennington-socher-manning:2014:EMNLP2014} in all neural taggers
and parsers.\footnote{\url{http://nlp.stanford.edu/data/glove.twitter.27B.zip}}

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & F1 \\
		\hline
		Stanford CoreNLP & 92.1 \\
		our bi-LSTM tokenizer (\S\ref{sec:tok}) & 93.6 \\
		\hline
	\end{tabular}
	\caption{\citet{owoputi-EtAl:2013:NAACL-HLT} POS tagging performance with automatic tokenization on
          the \textsc{Tweebank v2} test set. \label{tbl:pos-result-vs-tok}}
\end{table}

@@

Social media applications typically require processing large volumes
of data, making speed an important consideration. We therefore 
begin with the neural greedy stack LSTM parser introduced by \newcite{ballesteros-EtAl:2016:EMNLP2016},
which can parse a sentence in linear time and harnesses 
character representations, which should help mitigate the challenge of
spelling variation. We encourage the reader to refer their paper for
more details about the model.

@@

\paragraph{Ensemble.}  Due to ambiguity in the training
data---which most loss functions are not robust to \citep{Frnay2014ClassificationIT}, including the log loss we use, following
\citet{ballesteros-dyer-smith:2015:EMNLP}---and due to the instability
of neural network training, we follow \citet{Dietterich2000} and
consider an ensemble of twenty parsers trained using different random
initialization.  To parse at test time, the transition probabilities of the twenty
members of the ensemble are averaged.  The result achieves LAS of
79.4, outperforming all three systems above (Table~\ref{tbl:parse-result}).
However, ensembling 20 parsers also significantly slows down the parsing
speed and leads to the slowest system in our comparison.

\paragraph{Distillation.}  The shortcoming of the 20-parser ensemble is, of
course, that it requires twenty times the runtime of a single greedy
parser.  \newcite{kuncoro-16} proposed the distillation of 20
greedy transition-based parser into a single \emph{graph-based}
parser; they transformed the votes of the ensemble into a structured
loss function.  However, as Kuncoro et al.~pointed out, 
it is not straightforward to use a structured
loss in a \emph{transition-based} parsing algorithm.  Because fast
runtime is so important for NLP on social media, we
introduce a new way to distill our greedy ensemble into a single
transition-based parser (the first such attempt, to our knowledge).

>>>>

\section{Introduction}
NLP for social media messages is challenging, because domain
adaptation is required, and also because creating annotated datasets
(e.g., treebanks)
for training and evaluation is hard. 
Pioneering work by \citet{AAAIW113912} 
annotated 7,630 tokens' worth of tweets according to the
phrase-structure conventions of the Penn Treebank
\citep[PTB;][]{Marcus93buildinga}, enabling conversion to Stanford dependencies.
\citet{kong-EtAl:2014:EMNLP2014} further studied the challenges in
annotating tweets and presented a tweet treebank ({\sc Tweebank}),
consisting of 12,149 tokens and largely following conventions
suggested by \citet{schneider-EtAl:2013:LAW7-ID}, fairly close to 
\citet{Yamada03statisticaldependency} dependencies (without labels). 
Both annotation efforts were highly influenced by the PTB, whose guidelines
have good grammatical coverage on newswire. However, when it comes
to informal, unedited, user-generated text, the guidelines may leave
many annotation decisions unspecified.
%Motivation notes:  explain why UD is attractive (and more attractive
%than what previous work used, which was based on YM).  one reason is
%that it might allow sharing of texts annotated in different genres (or
%even languages, cite Waleed).  but universal annotation schemes are
%challenging!  this paper contends with the challenges of
%applying UD to social media texts.

@@

Based on these annotations, we nonetheless designed a pipeline to parse 
raw tweets into universal dependencies. Our pipeline includes: a
bidirectional LSTM (bi-LSTM) tokenizer, a word cluster-enhanced POS
tagger \citep[following][]{owoputi-EtAl:2013:NAACL-HLT}, and a stack LSTM parser
with character-based word representations
\cite{ballesteros-dyer-smith:2015:EMNLP}, which we refer to as our
``baseline'' parser.
To overcome the noise in our annotated data and achieve better performance
without sacrificing computational efficiency, we 
distill a 20-parser ensemble into a single greedy  parser 
\cite{DBLP:journals/corr/HintonVD15}.
We show further  that learning directly from the exploration of the ensemble parser
is more beneficial than learning from the gold standard ``oracle''
transition sequence. Experimental results show that an improvement of more
than 2.6 points in LAS over the
baseline parser can be achieved with our distillation method.  It outperforms
other state-of-the-art parsers in both accuracy and speed.

The contributions of this paper include:
\begin{itemize}
\item We study the challenges of annotating tweets in UD (\S\ref{sec:anno})
and create a new tweet treebank ({\sc Tweebank v2}), which includes 
tokenization, part-of-speech tagging, and labeled universal dependencies.
We also characterize the difficulties of creating such annotation.

@@

We first review {\sc Tweebank v1} of \citet{kong-EtAl:2014:EMNLP2014},
the previous largest Twitter dependency annotation effort 
(\S\ref{sec:tweebank}).
Then we introduce the differences in our tokenization
(\S\ref{sec:tok-anno}) and part-of-speech (\S\ref{sec:pos-anno}) (re)annotation with \citet{ICWSM101540} and 
\citet{gimpel-EtAl:2011:ACL-HLT2011}, respectively, on which {\sc Tweebank v1} was built. 
We describe our effort of adapting the
UD conventions to cover tweet-specific constructions (\S\ref{sec:ud-tweet}). 
Finally, we present our process of creating a new tweet treebank, {\sc
  Tweebank v2}, and characterize
the difficulties in reaching consistent annotations (\S\ref{sec:anno-process}).

@@

In the different direction of splitting tokens, UD guidelines also suggest to merge
{\it multi-token words} (e.g., {\it 20 000}) into one single token in some special
cases. We witnessed a small number of tweets that contain multi-token words
(e.g., {\it Y O}, and {\it R E T W E E T}) but didn't combine them for simplification.
Such tokens only account for 0.067\% and we use the UD
  {\it goeswith} relation to resolve these cases in the dependency annotations.

@@

Another important difference follows from the difference in
tokenization.  As discussed in \S\ref{sec:tok-anno}, UD calls for more
aggressive tokenization than that of \citet{ICWSM101540} which opted
out of splitting contractions and possessives. As a consequence of adopting \citet{ICWSM101540}'s tokenization, 
Gimpel et al.~introduced new parts of speech for these cases
instead.\footnote{These tags only account for 2.7\% of tokens,
  leading to concerns about data sparseness in tagging and all
  downstream analyses.}  For us, these tokens must be split, but
universal parts of speech can be applied.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth,trim={0 0 11cm 9cm},clip]{graphics/bilstm_tokenizer}
	\caption{The bi-LSTM tokenizer that segments `{\it its gonna be}' into `{\it it s gon na be}'.}\label{fig:tok-model}
\end{figure}
We introduce a new
character-level bidirectional LSTM (bi-LSTM) sequence-labeling model
\cite{DBLP:journals/corr/HuangXY15,ma-hovy:2016:P16-1}
for tokenization.
Our model takes the raw sentence and tags each character in this 
sentence as whether it is the beginning of a word (1 as the beginning and 0 otherwise).
Figure \ref{fig:tok-model} shows our tokenization model.
Space is treated as an input but deterministically \nascomment{check
  here}  assigned a special tag \$.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UD pipe v1.2 & 94.7 \\
		\hdashline
		UD pipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}

@@

Part-of-speech tagging for tweets has been extensively studied \cite{ritter-EtAl:2011:EMNLP,gimpel-EtAl:2011:ACL-HLT2011,derczynski-EtAl:2013:RANLP-2013,owoputi-EtAl:2013:NAACL-HLT,gui-EtAl:2017:EMNLP20172}. We therefore consider existing POS taggers for tweets instead of developing our own.
On the annotation scheme designed in \S\ref{sec:pos-anno}, based on UD and adapted for
Twitter, we compared several existing systems: the Stanford CoreNLP tagger, 
\citet{owoputi-EtAl:2013:NAACL-HLT}'s word cluster enhanced tagger
(both greedy and CRF variants), and 
\citet{ma-hovy:2016:P16-1}'s neural network POS tagger 
which has the state-of-the-art performance  on Penn Treebank data.
 %and also two bi-LSTM systems similar
%to \citet{DBLP:journals/corr/HuangXY15} and \citet{lample-EtAl:2016:N16-1}, one using
%word vectors and the other using word and character vectors (see Figure \ref{fig:pos-model}).
All systems were re-trained on the combination of the UD\_English and 
\textsc{Tweebank v2} training sets. We use Twitter-specific Glove embeddings released by
\citet{pennington-socher-manning:2014:EMNLP2014} in all neural taggers
and parsers.\footnote{\url{http://nlp.stanford.edu/data/glove.twitter.27B.zip}}

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it Tokenization System & $F_1$ \\
		\hline
		Stanford CoreNLP & 92.1 \\
		our bi-LSTM tokenizer (\S\ref{sec:tok}) & 93.6 \\
		\hline
	\end{tabular}
	\caption{\citet{owoputi-EtAl:2013:NAACL-HLT} POS tagging performance with automatic tokenization on
          the \textsc{Tweebank v2} test set. \label{tbl:pos-result-vs-tok}}
\end{table}

@@

Social media applications typically require processing large volumes
of data, making speed an important consideration. We therefore 
begin with the neural greedy stack LSTM parser introduced by \newcite{ballesteros-EtAl:2016:EMNLP2016},
which can parse a sentence in linear time and harnesses 
character representations to construct word vectors, which should help mitigate the challenge of
spelling variation. We encourage the reader to refer their paper for
more details about the model.

@@

\paragraph{Ensemble.}  Due to ambiguity in the training
data---which most loss functions are not robust to \citep{Frnay2014ClassificationIT}, including the log loss we use, following
\citet{ballesteros-dyer-smith:2015:EMNLP}---and due to the instability
of neural network training, we follow \citet{Dietterich2000} and
consider an ensemble of twenty parsers trained using different random
initialization.  To parse at test time, the transition probabilities of the twenty
members of the ensemble are averaged.  The result achieves LAS of
79.4, outperforming all three systems above (Table~\ref{tbl:parse-result}).

\paragraph{Distillation.}  The shortcoming of the 20-parser ensemble is, of
course, that it requires twenty times the runtime of a single greedy
parser, making it the slowest system in our comparison.  \newcite{kuncoro-16} proposed the distillation of 20
greedy transition-based parser into a single \emph{graph-based}
parser; they transformed the votes of the ensemble into a structured
loss function.  However, as Kuncoro et al.~pointed out, 
it is not straightforward to use a structured
loss in a \emph{transition-based} parsing algorithm.  Because fast
runtime is so important for NLP on social media, we
introduce a new way to distill our greedy ensemble into a single
transition-based parser (the first such attempt, to our knowledge).
